---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Overview

This report outlines metrics gathered from two months of activity data gathered from 
an anonymous individual collected during the months of October and November, 2012.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data


```{r load_data}
library(ggplot2)
all_activities <- read.csv(unzip("activity.zip", "activity.csv"))
file.remove("activity.csv")
activity <- all_activities[!is.na(all_activities$steps),]
activity$date <- as.Date(activity$date, "%Y-%m-%d")
summary(activity)
```

## What is mean total number of steps taken per day?

```{r total_steps}
mean(activity$steps, na.rm = TRUE )
```

```{r steps_per_day}
# summarize the steps by day
steps_per_day <- aggregate(list(steps = activity$steps), list(date = activity$date), FUN=sum)
gs <- ggplot(steps_per_day)
gs + geom_histogram(aes(x = steps), fill = "blue", color="green", bins =25) + labs(x = "steps per day")
summary(steps_per_day$steps)
```

## What is the average daily activity pattern?

```{r daily_activity_pattern}
average_steps_per_day <- aggregate(list(steps = activity$steps), list(date = activity$date), FUN=mean)
ga <- ggplot(data = average_steps_per_day)
ga + geom_line(aes(x = date, y = steps), color = "blue") + 
  geom_hline(yintercept = mean(average_steps_per_day$steps), linetype = "dotted")
activity[activity$steps == max(activity$steps, na.rm = TRUE), ]
```

## Imputing missing values

```{r missing values}
nrow(all_activities[is.na(all_activities$steps),])

# replace all NA's with zero and store it in a new data frame
no_nas <- all_activities
head(no_nas)
no_nas[is.na(no_nas)] <- mean(no_nas$steps, na.rm = TRUE)
head(no_nas)
no_nas_per_day <- aggregate(list(steps = no_nas$steps), list(date = no_nas$date), FUN=sum)
gn <- ggplot(data = no_nas_per_day)
gn + geom_histogram(aes(x = steps), fill = "blue", color="red", bins =25) + labs(x = "steps per day")
summary(no_nas_per_day$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r plot_weekends}
activity$weekday <- as.character(lapply(strftime(activity$date, format = "%w"), 
                           FUN = function(x) { if (x >= 1 && x <= 5) c("weekday") else c("weekend")}))
average_steps_per_interval <- aggregate(list(ave_steps = activity$steps), list(interval = activity$interval, weekday = activity$weekday), FUN=mean)

summary(average_steps_per_interval)
table(average_steps_per_interval$weekday)

gw <- ggplot(data = average_steps_per_interval, aes(x = interval, y = ave_steps, color=weekday))
gw + geom_line() + facet_grid(weekday ~ .)
```
